### 2025-04-22 paper/paper.tex

\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\linewidth]{SUMO_RUNNING.png}
    \caption{Example of SUMO running in through a GUI}
\end{figure}

@@

>>>>

\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\linewidth]{SUMO_RUNNING.png}
    \caption{Example of SUMO running in through SUMO-GUI}
\end{figure}

@@

our neural network is implemented through the python libraries traffic control interface (traci) and pytorch. Traci allows us to modify our simulation and get state information within a current timestep, while pytorch allows us to easily build and implement neural networks for our project.

The first step in designing the proposed neural network is to establish what an agent is, and how they will communicate with the model. Based on existing codebases and research, we decided implementing a decentralized multi agent reinforcement learning model. This specific implementation lacks a central coordinator, and each agent acts on its own in coordination with other agents. The agents were designed to represent an intersection with a traffic signal controller, and could accept the state of the lanes leading to the intersection, the state of the signals, and the states of vehicles at each intersection such as stopped time; it also accepted a list of possible actions the traffic signal controller could perform, which when provided to a deep-q learning model can guess the optimal move.

After designing an agent we decided to model the simulation environment in python to actually obtain states and actions for each agent. To do this, we used traci to obtain the states of vehicles on the lanes of traffic heading towards a specific traffic light, and assigned them to the traffic light. We then grabbed the possible state of actions, again assigning it to said traffic signal. This is neccessary for computing optimal actions later. We gather these every 5 seconds in the simulation, which greatly increases performance, and prevents rapidly switching signals during learning.

Our intitial plan was to predict the optimal states of every intersection with a traffic signal controller in Starkville, MS, however, upon implementation we realized our available hardware and timeframe would not allow us to do this. Instead, we opted to only use what we decided were points of interest, or intersections where traffic is particularly backed up. Most of these points were decided by considering the importance of the road, the time spent at intersections, and personal experience. This way we don't waste computational resources at intersections that see little traffic, or do fine directing traffic on their own. Doing so significantly decreases the computational requirements to run the model, and also allows us to focus only on important points.

Our model does not have preset correct moves, which makes designing a neural network a very difficult task. Therefore, its important to choose a (non supervised), q-score, action, learn.

implementing benchmarks
